---
title: "Primary energy consumption requirements based on SSPs"
author: "Marcus J Thomson"
date: "5/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(stringr)
library(scatterpie)
library(broom)
library(scales)
```

## Introduction

In this document, I compute the primary energy consumption (in MJ/day) for the global population from the present to 2100 based on five shared socioeconomic pathways (SSPs).

## Load data

### Population data

Read in population data from the SSP database at IIASA. Reference: [Keywan Riahi, et al. (2017). The Shared Socioeconomic Pathways and their energy, land use, and greenhouse gas emissions implications: An overview, Global Environmental Change, Volume 42, Pages 153-168, ISSN 0959-3780, DOI:110.1016/j.gloenvcha.2016.05.009](https://secure.iiasa.ac.at/web-apps/ene/SspDb/dsd?Action=htmlpage&page=10)

```{r}

output_path <- "/home/thomson/Projects/iep_food/IEP_food_demand_files/"
load(file = paste0(output_path, "ssp_world_long.Rdata"))

```


### Primary energy consumption

Read in the PEC from The Shift Project data portal: [TSP data portal (2020)](https://theshiftdataportal.org/).

The good people at The Shift Project have collected data from a variety of sources, and put these into online, open-access databases. I have pulled these data for PEC and organised it into a data frame, PEC_wide, containing energy consumption (in units of megatonnes oil equivalent, Mtoe) for each country, for years starting roughly from 1980 (some later). There are gaps in these data for some countries not covered, or some years missed, that I will need to fill later.

Among the data files are a few whose names defeat my (low level ninja) grep skills. Awkward column names were replaced, as was a strange time-stamp convention they're using for date. The new data frame, PEC_wide, is clean.

```{r}

path_main <- "~/Projects/iep_food/data/energy/TSP"
path_tsp  <- paste0(path_main, "/countries")

path_dir <- dir(path_tsp, 
                full.names = TRUE)

tmp1 <- path_dir[grep("Primary Energy Consumption by source, ", path_dir)]
tmp2 <- gsub("/home/thomson/Projects/iep_food/data/energy/TSP/countries/Primary Energy Consumption by source, ", "", tmp1)
tmp3 <- unlist(str_split(tmp2, ","))
tmp4 <- tmp3[order(tmp3)]
tmp5 <- c(" 1980-1991 (in Mtoe).csv", " 1980-1992 (in Mtoe).csv", " 1980-2016 (in Mtoe).csv", " 1986-2016 (in Mtoe).csv", " 1990-2016 (in Mtoe).csv", " 1992-2016 (in Mtoe).csv", " 1993-2016 (in Mtoe).csv", " 1994-2016 (in Mtoe).csv", " 1997-2016 (in Mtoe).csv", " 2003-2016 (in Mtoe).csv", " 2006-2016 (in Mtoe).csv", " 2008-2016 (in Mtoe).csv", " 2012-2016 (in Mtoe).csv")
tsp_countries <- as.data.frame(tmp4[!tmp4 %in% tmp5], 
                               stringsAsFactors = FALSE)
names(tsp_countries) <- "tsp_country"

```

I made separate files to: (1) map the country names used by TSP onto ISO-alpha2 and ISO-alpha3 codes as well as standard names from the FAO; and (2) group countries by geographic reigon. These files will appear separately in the repository.

```{r, echo = FALSE}

country_names <- read.csv(file = path_dir[grep("tsp_fao_country_mapper", path_dir)], 
                          header = TRUE, 
                          stringsAsFactors = FALSE)

fao_countrygroup <- read.csv(file = "~/Projects/iep_food/data/FAO/fao_countrygroup.csv", 
                             header = TRUE, 
                             stringsAsFactors = FALSE) %>% 
  dplyr::select(Country.Group, ISO3.Code) %>% 
  ungroup() %>% 
  rename(iso3 = ISO3.Code) %>% 
  filter(Country.Group %in% c("Southern Asia", "Southern Europe", "Northern Africa", "Oceania", "Middle Africa", "South America", "Western Asia", "Western Europe", "Eastern Europe", "Central America", "Western Africa", "Northern America", "Southern Africa", "South-eastern Asia", "Eastern Africa", "Northern Europe", "Eastern Asia", "Central Asia", "Caribbean"))

```

Since I will be making maps, I will also load map data included with the ggplot2 package. These data include some non-standard country names, so I made a mapping file for these as well.

```{r, echo = FALSE}
worldmap <- map_data("world") %>% 
  filter(region != "Antarctica")

mapdata_fao_mapper <- read.csv(file = "~/Projects/iep_food/data/FAO/mapdata_fao_mapper.csv", 
                               header = TRUE, 
                               stringsAsFactors = FALSE) %>% 
  rename(region = ggplot_mapdata, 
         iso3 = iso_code) %>% 
  full_join(fao_countrygroup, 
            by = "iso3")

```



```{r, echo = FALSE}

index <- grep("Primary Energy Consumption by source, ", 
              path_dir)

for(i in seq_along(index)){

  # tmp0 <- path_dir[grep(paste0("Primary Energy Consumption by source, ",
  #                              country_names[i, 1],
  #                              "\\b"),
  #                       path_dir)]
  tmp0 <- path_dir[i]
  tmp1 <- read.csv(file = tmp0,
                 header = TRUE,
                 stringsAsFactors = FALSE) %>%
  rename(oil = Oil,
         coal = Coal,
         gas = Gas,
         nuclear = Nuclear,
         hydro = Hydroelectricity,
         biomass_waste = Biomass.and.Waste,
         wind = Wind,
         solar_tidal_wave_fuelcell = Solar..Tide..Wave..Fuel.Cell,
         fuel_ethanol = Fuel.Ethanol,
         geothermal = Geothermal,
         biodiesel = Biodiesel,
         peat = Peat) %>%
  mutate(year = gsub("-01-01 00:00:00", " ", X) %>%
           as.integer()) %>%
  mutate(country     = country_names[i, 1],
         country_fao = country_names[i, 4],
         iso2        = country_names[i, 2],
         iso3        = country_names[i, 3]) %>%
  dplyr::select(country,
                iso3,
                year,
                oil:peat)

  if(i==1) Tmp <- tmp1
  if(i >1) Tmp <- rbind(Tmp, tmp1)

}

PEC_wide <- Tmp

# save(PEC_wide, file = "~/Projects/IEP_2/sourcefiles/global_pec_wide.Rdata")
```

NB: I think that these values are about 10 times higher than they should be. When I compare them to the per capita total energy from the World Bank data (and other sources ad hoc), keeping in mind that total energy consumption per capita should be above PEC per capita, I get roughly 10+ times below the PEC numbers. I will introduce a fudge factor below to rectify this -- but obviously I should understand where this difference comes from. (I have checked the Mtoe to MWh conversion.)

```{r, echo = FALSE}

load(file = "~/Projects/IEP_2/sourcefiles/global_pec_wide.Rdata") 

PEC_wide <- PEC_wide %>% 
  mutate(across(.cols = oil:peat, ~.x/10))

```


```{r}

loc_map <- readr::read_csv(file = "~/Data/IEP_Model/data_iep_model/location_code_map_data.csv", 
  col_types = cols(
    Location   = col_character(), 
    LocID      = col_number(), 
    ISO3_Code  = col_character(), 
    SubRegID   = col_number(), 
    SubRegName = col_character(), 
    Area.Code  = col_number(), 
    Area       = col_character())
) %>% 
  dplyr::select(-X1) %>% 
  drop_na()

country_regions <- loc_map %>% 
  dplyr::select(-LocID, -SubRegID, -Area.Code, -Area) %>% 
  dplyr::rename(
    REGION = ISO3_Code
  )

missing_locs <- data.frame(
  REGION = c("FSM", "CIV", "PRK", "PSE", "REU"), 
  Location = c("Micronesia, Federated States of", "Côte d'Ivoire", "Korea, Dem. Peop. Rep.", "Palestinian Territory", "Réunion"), 
  LocID = c(583, 384, 408, 275, 638), 
  SubRegName = c("Micronesia", "Western Africa", "Eastern Asia", "Western Asia", "Eastern Africa")
)

country_regions <- country_regions %>% 
  dplyr::bind_rows(missing_locs %>% 
                     dplyr::select(Location, REGION, SubRegName)) %>% 
  dplyr::rename(iso3 = REGION)



```

Fuel energy values are in Mtoe. Convert these to more sensible units (1 Mtoe = 1.163e10 kWh = 1.163e7 MWh): MWh. 

```{r}

conversion = 1.163e7

pec_countries <- PEC_wide %>% 
  dplyr::full_join(country_regions, 
                   by = "iso3") %>% 
  dplyr::filter(!is.na(SubRegName) & year >= 2000) %>% 
  dplyr::mutate(
    biodiesel = ifelse(!is.na(biodiesel), conversion*biodiesel, 0), 
    biomass_waste = ifelse(!is.na(biomass_waste), conversion*biomass_waste, 0), 
    coal = ifelse(!is.na(coal), conversion*coal, 0), 
    fuel_ethanol = ifelse(!is.na(fuel_ethanol), conversion*fuel_ethanol, 0), 
    gas = ifelse(!is.na(gas), conversion*gas, 0), 
    geothermal = ifelse(!is.na(geothermal), conversion*geothermal, 0), 
    hydro = ifelse(!is.na(hydro), conversion*hydro, 0), 
    nuclear = ifelse(!is.na(nuclear), conversion*nuclear, 0), 
    oil = ifelse(!is.na(oil), conversion*oil, 0), 
    peat = ifelse(!is.na(peat), conversion*peat, 0), 
    solar_tidal_wave_fuelcell = ifelse(!is.na(solar_tidal_wave_fuelcell), conversion*solar_tidal_wave_fuelcell, 0), 
    wind = ifelse(!is.na(wind), conversion*wind, 0) 
  ) %>% 
  dplyr::mutate(total = biodiesel + biomass_waste + coal + fuel_ethanol + gas + geothermal + hydro + nuclear + oil + peat + solar_tidal_wave_fuelcell + wind)

```



```{r}
pec_regions <- pec_countries %>% 
  group_by(SubRegName, year) %>% 
  summarise(
    biodiesel = sum(biodiesel, na.rm = TRUE), 
    biomass_waste = sum(biomass_waste, na.rm = TRUE),  
    coal = sum(coal, na.rm = TRUE), 
    fuel_ethanol = sum(fuel_ethanol, na.rm = TRUE), 
    gas = sum(gas, na.rm = TRUE), 
    geothermal = sum(geothermal, na.rm = TRUE), 
    hydro = sum(hydro, na.rm = TRUE), 
    nuclear = sum(nuclear, na.rm = TRUE), 
    oil = sum(oil, na.rm = TRUE), 
    peat = sum(peat, na.rm = TRUE), 
    solar_tidal_wave_fuelcell = sum(solar_tidal_wave_fuelcell, na.rm = TRUE), 
    wind = sum(wind, na.rm = TRUE), 
    total = sum(total, na.rm = TRUE), 
    .groups = "drop"
    )
```

Now relative quantities -- should sum to total = 1.

```{r}

pec_regions_rel <- pec_regions %>% 
  dplyr::mutate(
    biodiesel = ifelse(total != 0, biodiesel/total, 0), 
    biomass_waste = ifelse(total != 0, biomass_waste/total, 0), 
    coal = ifelse(total != 0, coal/total, 0), 
    fuel_ethanol = ifelse(total != 0, fuel_ethanol/total, 0), 
    gas = ifelse(total != 0, gas/total, 0), 
    geothermal = ifelse(total != 0, geothermal/total, 0), 
    hydro = ifelse(total != 0, hydro/total, 0), 
    nuclear = ifelse(total != 0, nuclear/total, 0), 
    oil = ifelse(total != 0, oil/total, 0), 
    peat = ifelse(total != 0, peat/total, 0), 
    solar_tidal_wave_fuelcell = ifelse(total != 0, solar_tidal_wave_fuelcell/total, 0), 
    wind = ifelse(total != 0, wind/total, 0) 
  ) %>% 
  dplyr::mutate(total = biodiesel + biomass_waste + coal + fuel_ethanol + gas + geothermal + hydro + nuclear + oil + peat + solar_tidal_wave_fuelcell + wind)

```



```{r}

pec_regions_rel_long <- pec_regions_rel %>% 
  dplyr::select(-total) %>% 
  tidyr::pivot_longer(data = ., 
                      cols = biodiesel:wind, 
                      names_to = "fuel")

```


```{r}

labs <- c("biodiesel", "biomass & waste", "coal", "fuel-ethanol", "gas", "geothermal", "hydro", "nuclear", "oil", "peat", "solar & fuelcell", "wind")

vals <- c("biodiesel" = "#ff5585", "biomass_waste" = "#CF8530", "coal" = "#320A26", "fuel_ethanol" = "#27c26f", "gas" = "#7D562E", "geothermal" = "#811DE2", "hydro" = "#1E1DE2", "nuclear" = "#E21D77", "oil" = "#7D2E2E", "peat" = "#736f64", "solar_tidal_wave_fuelcell" = "#C9CF30", "wind" = "#1D7EE2")

p1_asia <- ggplot(data = pec_regions_rel_long %>% 
                    dplyr::filter(SubRegName %in% c("Central Asia", "Eastern Asia", "South-Eastern Asia", "Southern Asia", "Western Asia")), 
       aes(x = year, 
           y = value)) + 
  geom_area(aes(fill = fuel), 
            alpha = 0.8) + 
  scale_fill_manual(
    labels = labs, 
    values = vals
    ) + 
  facet_wrap(~SubRegName) + 
  theme_minimal_hgrid() + 
  theme(
    legend.title = element_blank(), 
    legend.position = "bottom", 
    strip.text.x = element_text(size = 8), 
    axis.text.x = element_text(size = 8, angle = 90), 
    axis.text.y = element_text(size = 6), 
    axis.title = element_blank()
  ) + 
  scale_y_continuous(labels = scales::percent) + 
  labs(
    title = "Relative contribution to PEC"
  )

p1_africa <- ggplot(data = pec_regions_rel_long %>% 
                    dplyr::filter(SubRegName %in% c("Eastern Africa", "Middle Africa", "Northern Africa", "Southern Africa", "Western Africa")), 
       aes(x = year, 
           y = value)) + 
  geom_area(aes(fill = fuel), 
            alpha = 0.8) + 
  scale_fill_manual(
    labels = labs, 
    values = vals
    ) + 
  facet_wrap(~SubRegName) + 
  theme_minimal_hgrid() + 
  theme(
    legend.title = element_blank(), 
    legend.position = "bottom", 
    strip.text.x = element_text(size = 8), 
    axis.text.x = element_text(size = 8, angle = 90), 
    axis.text.y = element_text(size = 6), 
    axis.title = element_blank()
  ) + 
  scale_y_continuous(labels = scales::percent) + 
  labs(
    title = "Relative contribution to PEC"
  )

p1_europe <- ggplot(data = pec_regions_rel_long %>% 
                    dplyr::filter(SubRegName %in% c("Eastern Europe", "Northern Europe", "Southern Europe", "Western Europe")), 
       aes(x = year, 
           y = value)) + 
  geom_area(aes(fill = fuel), 
            alpha = 0.8) + 
  scale_fill_manual(
    labels = labs, 
    values = vals
    ) + 
  facet_wrap(~SubRegName) + 
  theme_minimal_hgrid() + 
  theme(
    legend.title = element_blank(), 
    legend.position = "bottom", 
    strip.text.x = element_text(size = 8), 
    axis.text.x = element_text(size = 8, angle = 90), 
    axis.text.y = element_text(size = 6), 
    axis.title = element_blank()
  ) + 
  scale_y_continuous(labels = scales::percent) + 
  labs(
    title = "Relative contribution to PEC"
  )

p1_newworld <- ggplot(data = pec_regions_rel_long %>% 
                    dplyr::filter(SubRegName %in% c("Caribbean", "Northern America", "South America")), 
       aes(x = year, 
           y = value)) + 
  geom_area(aes(fill = fuel), 
            alpha = 0.8) + 
  scale_fill_manual(
    labels = labs, 
    values = vals
    ) + 
  facet_wrap(~SubRegName) + 
  theme_minimal_hgrid() + 
  theme(
    legend.title = element_blank(), 
    legend.position = "bottom", 
    strip.text.x = element_text(size = 8), 
    axis.text.x = element_text(size = 8, angle = 90), 
    axis.text.y = element_text(size = 6), 
    axis.title = element_blank()
  ) + 
  scale_y_continuous(labels = scales::percent) + 
  labs(
    title = "Relative contribution to PEC"
  )

p1_oceania <- ggplot(data = pec_regions_rel_long %>% 
                    dplyr::filter(SubRegName %in% c("Australia/New Zealand", "Melanesia", "Micronesia", "Polynesia")), 
       aes(x = year, 
           y = value)) + 
  geom_area(aes(fill = fuel), 
            alpha = 0.8) + 
  scale_fill_manual(
    labels = labs, 
    values = vals
    ) + 
  facet_wrap(~SubRegName) + 
  theme_minimal_hgrid() + 
  theme(
    legend.title = element_blank(), 
    legend.position = "bottom", 
    strip.text.x = element_text(size = 8), 
    axis.text.x = element_text(size = 8, angle = 90), 
    axis.text.y = element_text(size = 6), 
    axis.title = element_blank()
  ) + 
  scale_y_continuous(labels = scales::percent) + 
  labs(
    title = "Relative contribution to PEC"
  )

p1_asia
p1_africa
p1_europe
p1_newworld
p1_oceania

```

Start with a fossil fuel phase-out for all regions by 2050. This will require primary energy consumption (PEC) to be supplied by non-GHG producing sources by 2050, sufficient to supply the future energy demands for their populations. So let's consider how each of the SSP populations will influence demands, given average energy demands for X kWh per person per year for each region on average.

```{r, echo = FALSE}

# download.file(url = "https://api.worldbank.org/v2/en/indicator/EG.USE.ELEC.KH.PC?downloadformat=csv", destfile = "/home/thomson/Projects/iep_food/data/energy/world_bank/electric_power_consumption.csv")

wb_epc <- read_csv(file = paste0("/home/thomson/Projects/iep_food/data/energy/world_bank/", "API_EG.USE.ELEC.KH.PC_DS2_en_csv_v2_2256478.csv"), skip = 3)

```
Per capita electricity consumption in kWh.

```{r}

wb_epc_long <- wb_epc %>% 
  dplyr::select(`Country Name`, `Country Code`, `2000`:`2014`) %>% 
  dplyr::rename(iso3 = `Country Code`) %>% 
  tidyr::drop_na() %>% 
  tidyr::pivot_longer(., 
                      cols = `2000`:`2014`, 
                      names_to = "year", 
                      values_to = "kWh") %>% 
  dplyr::mutate(year = as.integer(year))

head(wb_epc_long)

```

Find regional averages.

```{r}

wb_epc_long_regions <- wb_epc_long %>% 
  dplyr::full_join(country_regions %>% 
  dplyr::select(-Location), 
  by = "iso3") %>% 
  tidyr::drop_na()

head(wb_epc_long_regions)

```

Plot median consumption by region.

```{r}

p2 <- ggplot(wb_epc_long_regions, 
             aes(x = SubRegName, 
                 y = kWh*1e-3)) + 
  geom_boxplot(fill = "purple", 
               varwidth = FALSE, 
               outlier.alpha = 0, 
               alpha = 0.8) 

dat <- ggplot_build(p2)$data[[1]]

p2 + geom_jitter(colour = "darkblue", 
                 alpha = 0.2, 
                 size = 0.2) + 
  geom_segment(data = dat, 
               aes(x = xmin, 
                   xend = xmax, 
                   y = middle, 
                   yend = middle), 
               colour = "firebrick", 
               size= 1.2) + 
  cowplot::theme_minimal_hgrid() + 
  theme(
    axis.text.x = element_text(angle = 90, size = 8, vjust = 0.05, hjust = 0.95), 
    axis.title.x = element_blank()
  ) + 
  labs(
    title = "Per capita energy consumption for all years, 2000-2014", 
    y = "MWh per capita"
  )

```

Let's start with a target of 8 MWh per person per year, which is close to the median for Western Europe. Rules are: anyone who is presently above 8 MWh per capita declines to that average by 2050; anyone who is below it is allowed to rise to, but not exceed, that level by 2050. Additional constraint is that countries can only electrify at a maximum rate.

First we need to bring in population numbers.

```{r}

ssp_world_totals <- ssp_world_long %>% 
  dplyr::group_by(SCENARIO, REGION, year) %>% 
  dplyr::summarise(population = sum(millions, na.rm = TRUE)*1e6, 
                   .groups = "drop") %>% 
  dplyr::ungroup() %>% 
  dplyr::mutate(scenario = ifelse(
    SCENARIO == "SSP1_v9_130115", 
    "SSP1", 
    ifelse(
      SCENARIO == "SSP2_v9_130115", 
      "SSP2", 
      ifelse(
        SCENARIO == "SSP3_v9_130115", 
        "SSP3", 
        ifelse(
          SCENARIO == "SSP4d_v9_130115", 
          "SSP4", 
          "SSP5"
        ))))) %>% 
  dplyr::rename(iso3 = REGION) %>% 
  dplyr::select(-SCENARIO)

```


```{r}
# write_csv(ssp_world_totals, file = "~/Projects/ssp_pop/data/ssp_world_totals.csv")
```


The SSP population numbers are given in 5 year intervals, whereas the PEC values are annual. I will reduce the PEC interval to 5 years also, but take the values for each fuel type as the averages of the 3 years centered on 2010 and 2015; i.e., value(2010) = mean(values(2009,2010,2011)) and value(2015) = mean(values(2014,2015,2016)).

```{r}
pec_countries_reduced <- pec_countries %>% 
  dplyr::filter(year %in% c(2009,2010,2011,2014,2015,2016)) %>% 
  dplyr::mutate(bin_year = ifelse(year %in% c(2009, 2010, 2011), 
                                  2010, 
                                  2015)) %>% 
  dplyr::group_by(Location, iso3, SubRegName, bin_year) %>% 
  dplyr::summarise(
    biodiesel = mean(biodiesel, na.rm = TRUE), 
    biomass_waste = mean(biomass_waste, na.rm = TRUE), 
    coal = mean(coal, na.rm = TRUE), 
    fuel_ethanol = mean(fuel_ethanol, na.rm = TRUE), 
    gas = mean(gas, na.rm = TRUE), 
    geothermal = mean(geothermal, na.rm = TRUE), 
    hydro = mean(hydro, na.rm = TRUE), 
    nuclear = mean(nuclear, na.rm = TRUE), 
    oil = mean(oil, na.rm = TRUE), 
    peat = mean(peat, na.rm = TRUE), 
    solar_tidal_wave_fuelcell = mean(solar_tidal_wave_fuelcell, na.rm = TRUE), 
    wind = mean(wind, na.rm = TRUE), 
    total = mean(total, na.rm = TRUE), 
    .groups = "drop"
    ) %>% 
  dplyr::rename(year = bin_year)

pec_countries_reduced
```


```{r}

pec_countries_reduced_ssp <- pec_countries_reduced %>% 
  dplyr::full_join(ssp_world_totals, 
                   by = c("iso3", "year")) %>% 
  dplyr::mutate(per_capita_MWh = total/population) %>% 
  tidyr::drop_na()

head(pec_countries_reduced_ssp)

```

```{r}

p3 <- ggplot(pec_countries_reduced_ssp, 
             aes(x = SubRegName, 
                 y = per_capita_MWh)) + 
  geom_boxplot(fill = "purple", 
               varwidth = FALSE, 
               outlier.alpha = 0, 
               alpha = 0.8) 

dat1 <- ggplot_build(p2)$data[[1]] #median values from World Bank data
dat2 <- ggplot_build(p3)$data[[1]] #median values from these (TSP & IIASA SSP) data

p3 + geom_jitter(colour = "darkblue", 
                 alpha = 0.2, 
                 size = 0.2) + 
   # geom_segment(data = dat1, 
   #             aes(x = xmin, 
   #                 xend = xmax, 
   #                 y = middle, 
   #                 yend = middle), 
   #             colour = "darkblue", 
   #             size= 1.2) + 
  geom_segment(data = dat2, 
               aes(x = xmin, 
                   xend = xmax, 
                   y = middle, 
                   yend = middle), 
               colour = "firebrick", 
               size= 1.2) + 
  cowplot::theme_minimal_hgrid() + 
  theme(
    axis.text.x = element_text(angle = 90, size = 8, vjust = 0.05, hjust = 0.95), 
    axis.title.x = element_blank()
  ) + 
  labs(
    title = "Per capita PEC, 2010 & 2015", 
    y = "MWh per capita"
  )

```

Phase out all fossil fuels (coal, gas, oil, and peat) by 2050. Make up the difference with "wedges" of net-zero fuel types, to be decided later.

```{r}

pec_netzero_sources <- pec_countries_reduced_ssp %>% 
  dplyr::mutate(netzero_sources = biodiesel + biomass_waste + fuel_ethanol + geothermal + hydro + nuclear + solar_tidal_wave_fuelcell + wind, 
                fossil_sources = total - netzero_sources) %>% 
  dplyr::select(Location, iso3, SubRegName, year, netzero_sources, fossil_sources, total, population, scenario, per_capita_MWh)

head(pec_netzero_sources)

```

Extrapolate values for 2020, based on points for 2010 and 2015. For this, I use `nest()` and `lm()` functions.

```{r}

pec_lm_predictions <- pec_netzero_sources %>% 
  group_by(Location, iso3, SubRegName) %>% 
  nest() %>% 
  dplyr::mutate(fossil_sources_lm = map(data, ~ lm(fossil_sources ~ year, data = .x)), 
                netzero_sources_lm = map(data, ~ lm(netzero_sources ~ year, data = .x)), 
                fossil_sources_pred = map(fossil_sources_lm, ~predict(., data.frame(year = 2020))), 
                netzero_sources_pred = map(netzero_sources_lm, ~predict(., data.frame(year = 2020)))) %>% 
  unnest(cols = c(netzero_sources_pred, fossil_sources_pred)) %>% 
  ungroup()

```

In a few cases (e.g., US Virgin Islands, Syria, Trinidad & Tobago) this method breaks, producing negative estimates of net-zero and fossil fuel PEC for 2020. Here, I just zero out these negative numbers. We can return to this later with better country-level data, if necessary.

```{r}

pec_2020 <- pec_lm_predictions %>% 
  dplyr::mutate(year = 2020, 
                netzero_sources = ifelse(netzero_sources_pred > 0 & !is.na(netzero_sources_pred), 
                                         netzero_sources_pred, 
                                         0), 
                fossil_sources = ifelse(fossil_sources_pred > 0 & !is.na(fossil_sources_pred), 
                                         fossil_sources_pred, 
                                         0), 
                total = netzero_sources + fossil_sources) %>% 
  dplyr::select(Location, iso3, SubRegName, year, netzero_sources, fossil_sources, total) 

```

Join the population data frame.

```{r}

pec_pop_ssp_2020 <- pec_2020 %>% 
  full_join(ssp_world_totals %>% 
              dplyr::filter(year == 2020), 
            by = c("iso3", "year")) %>% 
  dplyr::mutate(per_capita_MWh = total/population) %>% 
  tidyr::drop_na() %>% 
  ungroup()

head(pec_pop_ssp_2020)

```



Create a data frame for countries and regions for years between 2020 and 2050.

```{r}

dimension <- pec_netzero_sources %>% 
  dplyr::distinct(Location, iso3, SubRegName, scenario) %>% 
  dim()
len <- dimension[1]/5 #the number of countries

tmp <- pec_netzero_sources %>% 
  dplyr::distinct(Location, iso3, SubRegName, scenario)

pec_netzero_2025_2050 <- data.frame(
  Location = rep(tmp$Location, 6), 
  iso3 = rep(tmp$iso3, 6), 
  SubRegName = rep(tmp$SubRegName, 6), 
  year = rep(seq(2025, 2050, by = 5), len), 
  netzero_sources = NA, 
  fossil_sources = NA, 
  total = NA, 
  scenario = rep(tmp$scenario, 6)
)

head(pec_netzero_2025_2050)

```


For target year 2050, I set regional target per capita PEC values, and pass these to the 

```{r}

regional_maxima <- data.frame(
  SubRegName = c("Southern Asia", "Southern Europe", "Northern Africa", "Middle Africa", "South America", "Western Asia", "Caribbean", "Australia/New Zealand", "Western Europe", "Eastern Europe", "Central America", "Western Africa", "Southern Africa", "South-Eastern Asia", "Eastern Africa", "Northern America", "Eastern Asia", "Northern Europe", "Melanesia", "Polynesia", "Micronesia", "Central Asia"), 
  target_value = c(5, 7, 4, 4, 4, 5, 5, 8, 8, 6, 4, 4, 4, 5, 4, 10, 8, 9, 4, 4, 6, 5)
)

head(regional_maxima)

```


```{r}

pec_pop_ssp_2025_2050 <- pec_netzero_2025_2050 %>% 
  dplyr::full_join(regional_maxima, 
                   by = "SubRegName") %>% 
  dplyr::rename(per_capita_MWh = target_value)

head(pec_pop_ssp_2025_2050)

```



```{r}

ssp_world_totals_2050 <- ssp_world_totals %>% 
  dplyr::filter(year == 2050) %>% 
  dplyr::full_join(pec_pop_ssp_2025_2050, 
                   by = c("iso3", "year", "scenario")) %>% 
  dplyr::select(Location, iso3, SubRegName, year, netzero_sources, fossil_sources, total, population, scenario, per_capita_MWh) %>% 
  dplyr::mutate(total = population*per_capita_MWh, 
                fossil_sources = 0, 
                netzero_sources = total - fossil_sources) %>% 
  tidyr::drop_na()

head(ssp_world_totals_2050)

```


```{r}

years_interp = data.frame(year = seq(2020, 2050, by = 5))

pec_2020_2050 <- pec_pop_ssp_2020 %>% 
  dplyr::bind_rows(ssp_world_totals_2050) %>% 
  dplyr::select(Location, iso3, SubRegName, year, netzero_sources, fossil_sources, scenario) %>% 
  tidyr::drop_na() %>% 
  dplyr::group_by(Location, iso3, SubRegName, scenario) %>% 
  tidyr::nest() %>% 
  dplyr::mutate(fossil_sources_lm = map(data, ~ lm(fossil_sources ~ year, data = .x)), 
                netzero_sources_lm = map(data, ~ lm(netzero_sources ~ year, data = .x)), 
                fossil_sources_pred = map(fossil_sources_lm, ~predict(., years_interp)), 
                netzero_sources_pred = map(netzero_sources_lm, ~predict(., years_interp))) %>% 
  unnest(cols = c(netzero_sources_pred, fossil_sources_pred)) %>% 
  ungroup()

head(pec_2020_2050)

```

Add years back into this.

```{r}

dimension <- pec_2020_2050 %>%
  distinct(iso3) %>% 
  dim()

len <- dimension[1] #number of countries
scen <- 5 #number of scenarios

yr <- data.frame(
  year = rep(rep(seq(2020, 2050, by = 5), scen), len)
)

pec_ssp_2020_2050 <- pec_2020_2050 %>% 
  dplyr::bind_cols(yr) %>% 
  dplyr::mutate(netzero_sources = netzero_sources_pred, 
                fossil_sources = fossil_sources_pred, 
                total = netzero_sources, fossil_sources) %>% 
  dplyr::select(Location, iso3, SubRegName, year, netzero_sources, fossil_sources, total, scenario)

head(pec_ssp_2020_2050)

```
Join population data frame.

```{r}

pec_pop_ssp_2020_2050 <- pec_ssp_2020_2050 %>% 
  dplyr::full_join(ssp_world_totals, 
                   by = c("iso3", "year", "scenario")) %>% 
  dplyr::mutate(per_capita_MWh = total/population) %>% 
  dplyr::select(Location, iso3, SubRegName, year, netzero_sources, fossil_sources, total, population, scenario, per_capita_MWh) %>% 
  dplyr::filter(year <= 2050) %>% 
  tidyr::drop_na()

head(pec_pop_ssp_2020_2050)

```

Bind all of the PEC data frames.

```{r}

pec_lm <- pec_netzero_sources %>% 
  dplyr::bind_rows(pec_pop_ssp_2020) %>% 
  dplyr::bind_rows(pec_pop_ssp_2020_2050) %>% 
  dplyr::bind_rows(
    ssp_world_totals_2050 %>% 
      dplyr::filter(year == 2050)
  ) %>% 
  tidyr::drop_na()

```


Aggregate into regions. (Problem looks to be here, as if it's aggregatign too many energy numbers and too few pop numbers.)

```{r}

pec_lm_regions <- pec_lm %>% 
  dplyr::group_by(SubRegName, year, scenario) %>% 
  dplyr::summarise(netzero_sources = sum(netzero_sources, 
                                         na.rm = TRUE), 
                   fossil_sources = sum(fossil_sources, 
                                        na.rm = TRUE), 
                   total = sum(total, 
                               na.rm = TRUE), 
                   population = sum(population, 
                                    na.rm = TRUE), 
                   .groups = "drop") %>% 
  ungroup() %>% 
  dplyr::mutate(per_capita_MWh = total/population)

head(pec_lm_regions)

```


```{r}

pec_lm_diff <- pec_lm_regions %>% 
  dplyr::mutate(tmp_netzero = ifelse(year == 2010, 
                                     netzero_sources, 
                                     NA), 
                tmp_fossil = ifelse(year == 2010, 
                                    fossil_sources, 
                                    NA)) %>% 
  dplyr::group_by(SubRegName, scenario) %>% 
  tidyr::fill(tmp_netzero, tmp_fossil) %>% 
  ungroup() %>% 
  dplyr::mutate(netzero_diff = netzero_sources-tmp_netzero, 
                fossil_diff = fossil_sources-tmp_fossil) %>% 
  dplyr::select(SubRegName, year, netzero_diff, fossil_diff, scenario)

head(pec_lm_diff)

```
There's a conversion error somewhere. At present, total global PEC is 576 EJ.

Maybe it's this: 1 MWh = 3.6e-9 EJ.

```{r}

p4_asia <- ggplot(data = pec_lm_relative %>% 
         dplyr::filter(year == 2050 & SubRegName %in% c("Central Asia", "Eastern Asia", "South-Eastern Asia", "Southern Asia", "Western Asia")), 
       aes(x = as.factor(year), 
           y = netzero_diff*(3.6e-9))) + 
  geom_col(aes(fill = scenario), 
           position = "dodge", 
           alpha = 0.8) + 
  facet_wrap(~SubRegName, 
             scale = "free_y") + 
  cowplot::theme_minimal_hgrid() + 
  theme(
    legend.position = "none", 
    legend.title = element_blank(), 
    strip.text.x = element_text(size = 8), 
    axis.text.x = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.y = element_text(size = 6), 
    axis.title.y = element_text(size = 8), 
    plot.tag = element_text(size = 10)
  ) + 
  labs(
    # title = "Absolute growth in zero emissions energy demand for 2050 by SSP", 
    tag = "Asia", 
    y = "EJ"
  ) + 
  panel_border() 

p4_africa <- ggplot(data = pec_lm_relative %>% 
         dplyr::filter(year == 2050 & SubRegName %in% c("Eastern Africa", "Middle Africa", "Northern Africa", "Southern Africa", "Western Africa")), 
       aes(x = as.factor(year), 
           y = netzero_diff*(3.6e-9))) + 
  geom_col(aes(fill = scenario), 
           position = "dodge", 
           alpha = 0.8) + 
  facet_wrap(~SubRegName, 
             scale = "free_y") + 
  cowplot::theme_minimal_hgrid() + 
  theme(
    legend.position = "none", 
    legend.title = element_blank(), 
    strip.text.x = element_text(size = 8), 
    axis.text.x = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.y = element_text(size = 6), 
    axis.title.y = element_text(size = 8), 
    plot.tag = element_text(size = 10)
  ) + 
  labs(
    # title = "Absolute growth in zero emissions energy demand for 2050 by SSP", 
    tag = "Africa", 
    y = "EJ"
  ) + 
  panel_border() 

p4_europe <- ggplot(data = pec_lm_relative %>% 
         dplyr::filter(year == 2050 & SubRegName %in% c("Eastern Europe", "Northern Europe", "Southern Europe", "Western Europe")), 
       aes(x = as.factor(year), 
           y = netzero_diff*(3.6e-9))) + 
  geom_col(aes(fill = scenario), 
           position = "dodge", 
           alpha = 0.8) + 
  facet_wrap(~SubRegName, 
             scale = "free_y") + 
  cowplot::theme_minimal_hgrid() + 
  theme(
    legend.position = "none", 
    legend.title = element_blank(), 
    strip.text.x = element_text(size = 8), 
    axis.text.x = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.y = element_text(size = 6), 
    axis.title.y = element_text(size = 8), 
    plot.tag = element_text(size = 10)
  ) + 
  labs(
    # title = "Absolute growth in zero emissions energy demand for 2050 by SSP", 
    tag = "Europe", 
    y = "EJ"
  ) + 
  panel_border() 

p4_newworld <- ggplot(data = pec_lm_relative %>% 
         dplyr::filter(year == 2050 & SubRegName %in% c("Caribbean", "Northern America", "South America")), 
       aes(x = as.factor(year), 
           y = netzero_diff*(3.6e-9))) + 
  geom_col(aes(fill = scenario), 
           position = "dodge", 
           alpha = 0.8) + 
  facet_wrap(~SubRegName, 
             scale = "free_y") + 
  cowplot::theme_minimal_hgrid() + 
  theme(
    legend.position = "none", 
    legend.title = element_blank(), 
    strip.text.x = element_text(size = 8), 
    axis.text.x = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.y = element_text(size = 6), 
    axis.title.y = element_text(size = 8), 
    plot.tag = element_text(size = 10)
  ) + 
  labs(
    # title = "Absolute growth in zero emissions energy demand for 2050 by SSP", 
    tag = "New World", 
    y = "EJ"
  ) + 
  panel_border() 

p4_oceania <- ggplot(data = pec_lm_relative %>% 
         dplyr::filter(year == 2050 & SubRegName %in% c("Australia/New Zealand", "Melanesia", "Micronesia", "Polynesia")), 
       aes(x = as.factor(year), 
           y = netzero_diff*(3.6e-9))) + 
  geom_col(aes(fill = scenario), 
           position = "dodge", 
           alpha = 0.8) + 
  facet_wrap(~SubRegName, 
             scale = "free_y") + 
  cowplot::theme_minimal_hgrid() + 
  theme(
    legend.position = "none", 
    legend.title = element_blank(), 
    strip.text.x = element_text(size = 8), 
    axis.text.x = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.y = element_text(size = 6), 
    axis.title.y = element_text(size = 8), 
    plot.tag = element_text(size = 10)
  ) + 
  labs(
    # title = "Absolute growth in zero emissions energy demand for 2050 by SSP", 
    tag = "Oceania", 
    y = "EJ"
  ) + 
  panel_border() 

```


```{r}

for_legend <- ggplot(data = pec_lm_relative %>% 
         dplyr::filter(year == 2050 & SubRegName %in% c("Australia/New Zealand", "Melanesia", "Micronesia", "Polynesia")), 
       aes(x = as.factor(year), 
           y = netzero_diff*(3.6e-9))) + 
  geom_col(aes(fill = scenario), 
           position = "dodge", 
           alpha = 0.8) +
  theme(
    legend.direction = "horizontal", 
    legend.title = element_blank()
  )

legend <- cowplot::get_legend(for_legend)

p4 <- cowplot::plot_grid(
  p4_asia, 
  p4_africa, 
  p4_europe, 
  p4_newworld, 
  p4_oceania, 
  legend, 
  ncol = 2
  )

title <- ggdraw() + draw_label("Absolute growth in zero emissions energy demand by 2050, by SSP", fontface='bold')

plot_grid(title, p4, ncol=1, rel_heights=c(0.1, 1)) # rel_heights values control title margins


```

Same for the entire world.
```{r}

pec_lm_relative_world <- pec_lm_relative %>% 
  dplyr::group_by(year, scenario) %>% 
  dplyr::summarise(netzero_diff = sum(netzero_diff, na.rm = TRUE)*(3.6e-9), 
                   fossil_diff = sum(fossil_diff, na.rm = TRUE)*(3.6e-9), 
                   .groups = "drop") %>% 
   dplyr::filter(year %in% c(2025, 2050))

```


```{r}

p4_world <- ggplot(data = pec_lm_relative_world, 
       aes(x = as.factor(year), 
           y = netzero_diff)) + 
  geom_col(aes(fill = scenario), 
           position = "dodge", 
           alpha = 0.8) + 
  cowplot::theme_minimal_hgrid() + 
  theme(
    legend.position = "bottom", 
    legend.title = element_blank(), 
    strip.text.x = element_text(size = 8), 
    axis.text.y = element_text(size = 6), 
    axis.title.y = element_text(size = 8), 
    plot.tag = element_text(size = 10), 
    plot.title = element_text(size = 12)
  ) + 
  labs(
    title = "Absolute growth in zero emissions energy demand for 2050 by SSP", 
    tag = "World", 
    x = "year", 
    y = "EJ"
  ) + 
  panel_border() 

p4_world

```

## Supply

Next, I must determine how much of this new demand can be met by wind and solar; and if there is a deficit, how much this will be.

### Wind 

What is the maximum available supply of wind energy? I use the  (The data frame 'wind_power' is calculated in `iep_energy_wind.Rmd`.)

```{r}
wind_power <- read_csv(file = "~/Data/IEP_Model/data_iep_model/wind_power.csv")
head(wind_power)
```

### Solar 

Here I use data provided by the Solar Atlas and World Bank termed "practical" photovoltaic power potential.

From data in: [Suri,Marcel; Betak,Juraj; Rosina,Konstantin; Chrkavy,Daniel; Suriova,Nada; Cebecauer,Tomas; Caltik,Marek; Erdelyi,Branislav. 2020. Global Photovoltaic Power Potential by Country (English). Energy Sector Management Assistance Program (ESMAP) Washington, D.C. : World Bank Group.](http://documents.worldbank.org/curated/en/466331592817725242/Global-Photovoltaic-Power-Potential-by-Country)

What is the maximum available supply of solar energy? (I calculate 'pvout' in `iep_energy_solar.Rmd`.)

```{r}
pvout <- read_csv(file = "~/Data/IEP_Model/data_iep_model/practical_solar_potential.csv")
head(pvout)
```

Photovoltaic potential energy is vastly greater than demand in all countries inspected.

### Geothermal

emissions: 91 g CO2/kWh in Bloomfield, K.K., Moore, J.N. and Neilson, R.N., 2003. Geothermal energy reduces greenhouse gases. Geothermal Resources Council Bulletin, 32(2), pp.77-79.
