---
title: "Solar energy supply"
author: "Marcus J Thomson"
date: "5/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Solar 

Here I use data provided by the Solar Atlas and World Bank termed "practical" photovoltaic power potential.

From data in: [Suri,Marcel; Betak,Juraj; Rosina,Konstantin; Chrkavy,Daniel; Suriova,Nada; Cebecauer,Tomas; Caltik,Marek; Erdelyi,Branislav. 2020. Global Photovoltaic Power Potential by Country (English). Energy Sector Management Assistance Program (ESMAP) Washington, D.C. : World Bank Group.](http://documents.worldbank.org/curated/en/466331592817725242/Global-Photovoltaic-Power-Potential-by-Country)

What is the maximum available supply of solar energy?

```{r, echo=FALSE}

solar_potential_data <- read_csv(file = "~/Data/IEP_Model/data_iep_model/SolarGIS solar potential data - harmonized summary.csv", skip = 1)

```

The column labelled "Median" is the median theoretical potential (GHI) measured in kWh/m^2/day long-term. The column labelled "Median_1" is the median practical potential (PVOUT) for Level 1, measured in kWh/kWp/day long-term.

```{r}
practical_solar_potential <- solar_potential_data %>%
  dplyr::rename(
    "iso3" = "ISO_A3",
    "country" = "Country or region",
    "evaluated_area_km2" = "Evaluated area",
    "level_1_area_ratio" = "Level 1 area \n(% of evaluated area)",
    "minimum_pvout" = "Minimum_1",
    "x10th_percentile_pvout" = "10th percentile_1",
    "x25th_percentile_pvout" = "25th Percentile_1",
    "mean_pvout" = "Average_1",
    "median_pvout" = "Median_1",
    "x75th_percentile_pvout" = "75th Percentile_1",
    "x90th_percentile_pvout" = "90th percentile_1",
    "maximum_pvout" = "Maximum_1"
  ) %>% 
  dplyr::select(iso3, country, evaluated_area_km2, level_1_area_ratio, minimum_pvout, x10th_percentile_pvout, x25th_percentile_pvout, minimum_pvout, mean_pvout, median_pvout, x75th_percentile_pvout, x90th_percentile_pvout, maximum_pvout) %>% 
  dplyr::mutate(
    practical_pvout_MJ = evaluated_area_km2*level_1_area_ratio*mean_pvout*(1e6)*365*3.6, 
    practical_pvout_EJ = practical_pvout_MJ*1e-12
  )

head(practical_solar_potential)

```

```{r}
# write_csv(practical_solar_potential, file = "~/Data/IEP_Model/data_iep_model/practical_solar_potential.csv")
```

## Capacity

The real issue with solar potential is how rapidly solar collectors, such as photovoltaic power stations, can be distributed at scale. I need to estimate this.

```{r}

irena_data <- read_csv(file = "~/Data/IEP_Model/data_iep_model/IRENA_data - data.csv")

irena_data

```

